---
title: 'Lab assignment 1: Processing models'
author: "Luka van der Plas (4119142) & Rianne Lam (6888216)"
date: "15 november 2019"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section I: Human data

### Question 1:

A.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

load("keyPressDataWithLaneDeviation.Rdata")
  
NoErr <- subset(keyPressDataWithLaneDeviation, typingErrorMadeOnTrial == 0)
totalDialingTime <- subset(NoErr, phoneNrLengthAfterKeyPress == 11)

summaryStatistics <- function(data) {
  meanCond <- mean(data)
  sdCond <- sd(data)
  seCond <- sdCond / sqrt(length(data))
  return(c(meanCond, sdCond, seCond))
}

summarizeData <- function(dataToSummarize) {

totalDialingTimePP <- aggregate(dataToSummarize[ , c("pp", "timeRelativeToTrialStart")], by = list(dataToSummarize$pp), FUN = mean)$timeRelativeToTrialStart

summary <- summaryStatistics(totalDialingTimePP)

return(summary)
} 
  
summaryDataCond1 <- summarizeData(subset(totalDialingTime, partOfExperiment == "dualSteerFocus"))
summaryDataCond2 <- summarizeData(subset(totalDialingTime, partOfExperiment == "dualDialFocus"))

dialingTime <- c(summaryDataCond1, summaryDataCond2)

results2 <- data.frame("M" = paste(round(c(summaryDataCond1[1], summaryDataCond2[1]), 3),  " ms"),
                   "SD" = paste(round(c(summaryDataCond1[2], summaryDataCond2[2]), 3),  " ms"),
                   "SE" = paste(round(c(summaryDataCond1[3], summaryDataCond2[3]), 3),  " ms"))
rownames(results2) <- c("Steering focus", "Dialing focus")

results2

```

B.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

avg_dev_trial <- function(data, trialN) {
  dataTrial <- subset(data, data$trial == trialN)
  abs_lane_pos <- mean(abs(dataTrial$lanePosition), na.rm = TRUE)
  return(abs_lane_pos)
}

avg_dev_cond <- function(condition) {
  dataCond <- subset(NoErr, NoErr$partOfExperiment == condition) 
  ppCount <- max(dataCond$pp)
  abs_lane_pos <- vector()
  for (i in 1:ppCount) {
      dataCondpp <- subset(dataCond, dataCond$pp == i)
      trials <- unique(dataCondpp$trial)
    for (j in trials) {
      mean_abs_lane_pos <- avg_dev_trial(dataCondpp, j)
      abs_lane_pos <- append(abs_lane_pos, mean_abs_lane_pos)
    }
  }
  
  return(summaryStatistics(abs_lane_pos))
} 

steerFocus <- avg_dev_cond("dualSteerFocus")
dialFocus <- avg_dev_cond("dualDialFocus")

laneDev <- c(steerFocus, dialFocus)

results3 <- data.frame("M" = paste(round(c(steerFocus[1], dialFocus[1]), 3),  " m"),
                   "SD" = paste(round(c(steerFocus[2], dialFocus[2]), 3),  " m"),
                   "SE" = paste(round(c(steerFocus[3], dialFocus[3]), 3),  " m"))
rownames(results3) <- c("Steering focus", "Dialing focus")

results3


```

C.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

library(ggplot2)
library(dplyr)

mean_dev_pp_df <- NoErr %>%
  # Group by participant, keypress, condition
  group_by(pp, phoneNrLengthAfterKeyPress, partOfExperiment) %>%  
  # Take the mean of absolute lane position for these groups
  summarise(mean_dev_pp = mean(abs(lanePosition)))

mean_dev_all <- mean_dev_pp_df %>%
  # Group by keypress, condition
  group_by(phoneNrLengthAfterKeyPress, partOfExperiment) %>%
  # Take the mean of the mean for these groups per participant (mean across participants)
  summarise(mean_dev = mean(mean_dev_pp)) %>%
  # Only look at the two relevant conditions
  filter(partOfExperiment %in% c("dualDialFocus", "dualSteerFocus")) 

# Replace condition names for plot
mean_dev_all$partOfExperiment <- recode(mean_dev_all$partOfExperiment, "dualDialFocus" = "Dialing-focus", "dualSteerFocus" = "Steering-focus")

n <- max(unique(keyPressDataWithLaneDeviation$pp))

plotHumanData <- ggplot(mean_dev_all, aes(x = phoneNrLengthAfterKeyPress, y = mean_dev, group = partOfExperiment)) +
  geom_line(size = 1) +
  geom_point(aes(shape = factor(partOfExperiment)), size = 3.5) +
  theme_classic() +
  theme(legend.position = c(0.9, 0.1), legend.title = element_blank()) + 
  ylab("Lateral Deviation (m)") + 
  xlab("Number of key presses") + 
  geom_errorbar(aes(ymin = mean_dev - (sd(mean_dev) / sqrt(n)), ymax = mean_dev + (sd(mean_dev) / sqrt(n))), size = 1, width = 0.2, position = position_dodge(0.05)) +
  scale_x_continuous(breaks = seq(0, 12, 1)) 

plotHumanData

```

D. For both conditions, it seems that participants did not interleave dialing for driving between the 5th and 6th key presses. At the 6th key press, the absolute lane deviation did not decrease compared to the 5th key press, while this would have been expected if participants were to focus on the road at this point.

However, in the dialing focus condition we can see that from the 6th key press onwards participants tend to focus less on the road and thus show greater absolute lane deviation. This gives a kind of breakpoint in the plot. But this result cannot be explained by the hypothesis that particpants would interleave dialing for driving at this point in time. 

### Question 2

A. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
load("tableOfDriftValuesCalibration.Rdata")

driftData <- subset(tableOfDriftValuesCalibration, trialTime >= 15000 & trialTime <= 18000)
driftData$trial <- as.factor(driftData$trial)

ggplot(driftData, aes(trialTime, posX)) +
  geom_line(aes(group = trial, color = trial)) +
  theme(legend.position = "none")


```

B.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#get trials and timesteps
trials <- unique(tableOfDriftValuesCalibration$trial)
timesteps <- 3000 / 50
timerange <- seq(0, 3000, 50)

#empty df
trial_data <- data.frame()

#add data for trials
for (t in trials) {
  deviations <- rnorm(timesteps, mean = 0, sd = 0.13)
  deviations <- cumsum(deviations)
  
  for (i in 1:length(deviations)) {
    trial_data <- rbind(trial_data, c(t, timerange[i], deviations[i]))
  }
}

#set column names
colnames(trial_data) <- c("trial", "time", "deviation")
trial_data$trial <- as.factor(trial_data$trial)

ggplot(trial_data, aes(time, deviation)) +
  geom_line(aes(group = trial, color = trial)) +
  theme(legend.position = "none")

```

C. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}

min_pos <- min(min(driftData$posX), min(trial_data$deviation))
min_pos <- floor(min_pos * 10) / 10
max_pos <- max(max(driftData$posX), max(trial_data$deviation))
max_pos <- ceiling(max_pos * 10) / 10

ggplot(driftData, aes(posX)) +
  geom_histogram(binwidth = 0.2) +
  scale_x_continuous(breaks = seq(min_pos, max_pos, 0.2)) +
  coord_cartesian(xlim=  c(min_pos, max_pos), ylim=c(0, 350)) +
  xlab("lateral position") +
  ylab("frequency")

ggplot(trial_data, aes(deviation)) +
  geom_histogram(binwidth = 0.2) +
  scale_x_continuous(breaks = (breaks = seq(min_pos, max_pos, 0.2))) +
  coord_cartesian(xlim=  c(min_pos, max_pos), ylim=c(0, 350)) +
  xlab("lateral position") +
  ylab("frequency")

```

D.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
sdDf <- data.frame("Data Type" = c("Human Data", "Simulated Data"), "Standard Deviation" = c(paste(0.359, "ms"), paste(0.710, "ms")))
                   
sdDf

```

E.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# Function

trySds <- function(sdToTry) {
  
  #get trials and timesteps
  trials <- 50
  timesteps <- 3000 / 50
  timerange <- seq(0, 3000, 50)
  
  #empty df
  trial_data <- data.frame()
  
  #add data for trials
  for (t in trials) {
    deviations <- rnorm(timesteps, mean = 0, sd = sdToTry)
    deviations <- cumsum(deviations)
    
    for (i in 1:length(deviations)) {
      trial_data <- rbind(trial_data, c(t, timerange[i], deviations[i]))
    }
  }
  
  #set column names
  colnames(trial_data) <- c("trial", "time", "deviation")
  trial_data$trial <- as.factor(trial_data$trial)
  
  plot1 <- ggplot(trial_data, aes(time, deviation)) +
    geom_line(aes(group = trial, color = trial)) +
    theme(legend.position = "none")
  
  plot2 <- ggplot(trial_data, aes(deviation)) +
    geom_histogram(binwidth = 0.2) +
    xlab("lateral position") +
    ylab("frequency")
  
  sdDiff <- abs(0.359 - sd(trial_data$deviation))
  
  return(sdDiff)
}

sdsToTry <- seq(0, 0.13, 0.001)
results <- vector()

for(i in sdsToTry) {
  results <- append(results, trySds(i))
}

# which(results == min(results))
# results[99]
# sdsToTry[99]

```

### Question 3

We changed the value of gaussDeviateSD to 0.07 and we changed the value of gaussDriveNoiseSD to 0.053 so that the difference would be the same ratio as of the previous standard deviations.

A. The average value of the interkeypress intervals is 273 (rounded towards the nearest integer).
B. We used 275 as the interkeypress interval for the model because we consider this amount of rounding to be appropriate. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
source("DrivingModel.R")

dataSingleDialing <- subset(NoErr, partOfExperiment == "singleDialing2")

totalTrialTime <- dataSingleDialing %>%
  select(-9) %>%
  group_by(pp, trial) %>%
  filter(phoneNrLengthAfterKeyPress == max(phoneNrLengthAfterKeyPress) & Event1 != "Start")

meanInterval <- totalTrialTime$timeRelativeToTrialStart / totalTrialTime$phoneNrLengthAfterKeyPress
  
meanIntervalAll <- round(sum(meanInterval) / length(meanInterval), 0)

```

### Question 4

A. 

1 simulation:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results1 <- runAllSimpleStrategies(nrSimulations = 1)

```

5 simulations:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results5 <- runAllSimpleStrategies(nrSimulations = 5)

```

10 simulations:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results10 <- runAllSimpleStrategies(nrSimulations = 10)

```

50 simulations:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results50 <- runAllSimpleStrategies(nrSimulations = 50)

```

100 simulations:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results100 <- runAllSimpleStrategies(nrSimulations = 100)

```

200 simulations:

```{r, echo = FALSE, warning = FALSE, message = FALSE}

results200 <- runAllSimpleStrategies(nrSimulations = 200)

```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

data.frame("N of simulations"= c(1, 5, 10, 50, 100, 200), "Correlation trial time and lateral deviation"= c(results1, results5, results10, results50, results100, results200))

```

B. We can see that as the number of simulations increases, the data becomes less noisy and more stable. Additionally, as the number of simulations increases the range of the average lateral deviation decreases somewhat. Furthermore, we looked at the correlation between trial time and lateral deviation for all simulation lengths. The correlation becomes larger as the number of simulations increases, and clearly reaches a maximum at 100 simulations. This means that somewhere around 100 simulations is the optimal choice (data is stable an additional simulations provide no information gain anymore).
However, the computing time increases quite a bit as we reach 100-200 simulations. Therefore, we suggest that it would be best to use around 50 simulations when still working on the model, and then use around 100 simulations for the final model predictions. 

### Question 5

```{r, echo = FALSE, warning = FALSE, message = FALSE}

generate_strategies <- function(n) {
  #give all strategies of length n
  
  #base case
  if (n ==1) {
    return (matrix(c(FALSE, TRUE)))
  }
    
  #recursive definition
  tail <- generate_strategies(n - 1)
  count <- dim(tail)[1]
  
  start_0 <- cbind(matrix(rep(FALSE, count)), tail)
  start_1 <- cbind(matrix(rep(TRUE, count)), tail)
  
  return(rbind(start_0, start_1))
}

all_strategies <- generate_strategies(10)

give_breakpoints <- function(strat) {
  breakpoints <- c()
  
  for (i in 1:length(strat)) {
    if (strat[i]) {
      breakpoints <- c(breakpoints, c(i))
    }
  }
  breakpoints
}

steeringTimeOptions <- c(2,4,6,8,10,12)

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

runAllComplexStrategies <- function(nrSimulations,phoneNumber) {
	
	
	normalPhoneStructure <- c(1,6)  ### indicate at what digit positions a chunk needs to be retrieved (1st and 6th digit)
	phoneStringLength <- 11   ### how many digits does the number have?
	

	### vectors that will contain output of the simulation. These are later used to create 1 table with all values
	keypresses <- c()
	times <- c()
	deviations <- c()
	strats <- c()
	steers <- c()	

	### iterate through all strategies
	## in this simple model we assume that a participant uses a consistent strategy throughout the trial. That is, they only type each time 1 digit, or type 2 digits at a time, or type 3 digits at a time (i.e., all possible ways of 1:phoneStringLength: 1, 2,3,4, ...11)
	
	for (i in 1:nrow(all_strategies))
	{
		## quick way of calculating positions to interleave: repeat strategy & multiply with position in vector (e.g., 333*123 = 369 this means: you interleave BEFORE every 3rd digit (333), and there are 3 positions to interleave (1st, 2nd, 3rd, or 123). Therefore you interleave BEFORE digits 3 (3*1), 6 (3*2), and 9 (3*3))
	  
	  strategy <- give_breakpoints(all_strategies[i,])
		

		locSteerTimeOptions <- steeringTimeOptions
		if (length(strategy) == 0)
		{
			locSteerTimeOptions <- c(0)
		}



		### now run a trial (runOneTrial) for all combinations of how frequently you update the steering when you are steering (locSteerTimeOptions) and for the nuber of simulations that you want to run for each strategy (nrSimulations)
		for (steerTimes in locSteerTimeOptions)
		{
			for (j in 1:nrSimulations)
			{

				### run the simulation and store the output in a table
				locTab <- runOneTrial(strategy, steerTimes,normalPhoneStructure,phoneStringLength,phoneNumber)
	
				##only look at rows where there is a keypress
				locTab <- locTab[locTab$events == "keypress",]
		
				### add the relevant data points to variables that are stored in a final table
				keypresses <- c(keypresses,1:nrow(locTab))
				times <- c(times,locTab$times)
				deviations <- c(deviations,locTab$drifts)
				strats <- c(strats,rep(i,nrow(locTab)))
				steers <- c(steers,rep(steerTimes,nrow(locTab)))
		
			}
		}#end of for steerTimes	

	}##end of for nr strategies
	
	
	### now make a new table based on all the data that was collected
	tableAllSamples <- data.frame(keypresses,times,deviations,strats,steers)
	
	#### In the table we collected data for multiple simulations per strategy. Now we want to know the average performane of each strategy.
	#### These aspects are calculated using the "aggregate" function
	
	
	## calculate average deviation at each keypress (keypresses), for each unique strategy variation (strats and steers)
	agrResults <- with(tableAllSamples,aggregate(deviations,list(keypresses=keypresses, strats= strats, steers= steers),mean))
	agrResults$dev <- agrResults$x
	
	
	### also calculate the time interval
	agrResults$times <- with(tableAllSamples,aggregate(times,list(keypresses=keypresses, strats= strats, steers= steers),mean))$x
		
	
	###now calculate mean drift across the trial
	agrResultsMeanDrift <-  with(agrResults,aggregate(dev,list(strats= strats, steers= steers),mean))
	agrResultsMeanDrift$dev <- agrResultsMeanDrift$x
	
	### and mean trial time
	agrResultsMeanDrift$TrialTime <-  with(agrResults[agrResults$keypresses ==11,],aggregate(times,list( strats= strats, steers= steers),mean))$x	
	
	return(agrResultsMeanDrift)
}

```

The plot below shows the simulation results together with the human data

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# results <- runAllComplexStrategies(nrSimulations = 50)

# save(results, file = "ComplexStrategies50Sim.R")

load("ComplexStrategies50Sim.R")

results$TrialTime <- results$TrialTime / 1000

dialingTime <- dialingTime / 1000
dialTimeDialMean <- dialingTime[4]
dialTimeSteerMean <- dialingTime[1]
dialTimeDialSE <- dialingTime[6]
dialTimeSteerSE <- dialingTime[3]

laneDevDialMean <- laneDev[4]
laneDevSteerMean <- laneDev[1]
laneDevDialSE <- laneDev[6]
laneDevSteerSE <- laneDev[3]

#get trials that only interleave between 5 and 6

#find index of target strategy
target_strat <- c(FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE)
target_strat_n <- 0
for(i in 1:nrow(all_strategies)) {
  if (identical(all_strategies[i,], target_strat)) {
    target_strat_n <- i
    break
  }
}

natural_breakpoint_sims <- subset(results, strats == target_strat_n)

less_than_5_breakpoints <- c()

for (i in 1:nrow(results)) {
  stratIdx <- results[i, "strat"] 
  less_than_5_breakpoints[i] <- sum(all_strategies[stratIdx, ]) < 5
}

results <- cbind(results, less_than_5_breakpoints)

library(ggplot2)

#plot all strategies
ggplot(data = results) +
  geom_point(aes(TrialTime, abs(dev), color="Simulations"), size = 0.2) +
  
  geom_point(aes(x = dialTimeSteerMean, y = laneDevSteerMean, color = "Steer focus trials")) +
  geom_errorbar(aes(x = dialTimeSteerMean, ymin = laneDevSteerMean - laneDevSteerSE, ymax = laneDevSteerMean + laneDevSteerSE, color = "Steer focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = dialTimeSteerMean - dialTimeSteerSE, xmax = dialTimeSteerMean + dialTimeSteerSE, y = laneDevSteerMean, color = "Steer focus trials", height = 0.02)) +
  
  geom_point(aes(x = dialTimeDialMean , y = laneDevDialMean, color =  "Dial focus trials")) +
  geom_errorbar(aes(x = dialTimeDialMean , ymin = laneDevDialMean - laneDevDialSE, ymax = laneDevDialMean + laneDevDialSE, color =  "Dial focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = dialTimeDialMean  - dialTimeDialSE, xmax = dialTimeDialMean  + dialTimeDialSE, y = laneDevDialMean, color =  "Dial focus trials", height = 0.02)) + 
  
  geom_point(data = natural_breakpoint_sims, aes(TrialTime, abs(dev), color="Natural breakpoint simulations")) +
  
  
  xlab("Dial time (s)") +
  ylab("Average Lateral Deviation (m)") +
  labs(color = "Data source")


```

B.
In the dial-focus condition, performance seems to fall outside of the performance space of the model. Participants are very fast in dialing and show a relatively high lane deviation; therefore their performance is not very efficient.
In the steer-focus condition, performance does seem to fall on the outside of the edge of the performance space. Participants are fast in dialing and show intermediate lane deviation compared to the model simulations. Therefore, their performance seems to be efficient (fast but no ver high lane deviation). 

C.
We selected the simulations for which the data fall within the margin of error of the steering condition for the human trials. This gave 30 strategies. Below, the breakpoints for each of these strategies is displayed.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)

stratsDial <- results %>%
  filter(TrialTime > (dialTimeDialMean - dialTimeDialSE) & TrialTime < (dialTimeDialMean + dialingTime[3])) %>%
           filter(abs(dev) > (laneDevDialMean - laneDevDialSE) & abs(dev) < (laneDevDialMean + laneDevDialSE))

stratsSteer <- results %>%
  filter(TrialTime > (dialTimeSteerMean - dialTimeSteerSE) & TrialTime < (dialTimeSteerMean + dialTimeSteerSE)) %>%
           filter(abs(dev) > (laneDevSteerMean - laneDevSteerSE) & abs(dev) < (laneDevSteerMean + laneDevSteerSE)) 

break_data <- data.frame()

for (i in 1:nrow(stratsSteer)) {
  strat <- all_strategies[i, ]
  breaks <- which(strat)
  
  number_of_breaks <- length(breaks)
  first_break <- breaks[1]
  break_data <- rbind(strat, break_data)
}

```

```{r q5c, echo = FALSE}

within_steer_data <- subset(complex_sims_results, TrialTime >= steer_dial_m - steer_dial_se & TrialTime <= steer_dial_m + steer_dial_se)
within_steer_data <- subset(within_steer_data, dev >= steer_dev_m - steer_dev_se & dev <= steer_dev_m + steer_dev_se)

within_dial_data <- subset(complex_sims_results, TrialTime >= dial_dial_m - dial_dial_se & TrialTime <= dial_dial_m + dial_dial_se)
within_dial_data <- subset(within_dial_data, dev >= dial_dev_m - dial_dev_se & dev <= dial_dev_m + dial_dev_se)

#matching_strategies <- c(within_steer_data$strats, within_dial_data$strats)

matching_strategies <- within_steer_data$strats
matching_strategies <- sort(matching_strategies)

break_df <- data.frame()
break_matrix <- matrix(nrow = length(matching_strategies), ncol = 10)
plot_df <- data.frame()

for (i in 1:length(matching_strategies)) {
  index <- matching_strategies[i]
  strat <- all_strategies[index,]
  breaks <- which(strat)
  
  number_of_breaks <- length(breaks)
  first_break <- breaks[1]
  
  for (br in breaks) {
    break_df <- rbind(break_df, c(i, br))
  }
  
  break_matrix[i,] <- strat
  
  for (pos in 1:10) {
    plot_df <- rbind(plot_df, c(i, pos, strat[pos]))
  }
  
}

colnames(break_df) <- c("strategy", "break_position")
colnames(plot_df) <- c("strategy", "keypress", "interleave")

ggplot(plot_df) +
  geom_raster(aes(x = keypress, y = strategy, fill = factor(interleave))) +
  scale_fill_manual(values = c("cadetblue3", "coral3")) +
  scale_x_continuous(breaks = seq(1, 10, 1), minor_breaks = NULL) +
  scale_y_continuous(breaks = seq(1,length(matching_strategies),1), minor_breaks = NULL) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

```

A few things to note:
* The first breakpoint can be anywhere from the first to the tenth digit. However, there are slightly simulations where the first breakpoint was in digits 2-4.
* All of the strategies have only one or two breakpoints.
* There does not seem to be a bias towards any particular position for breakpoints (like the 5th digit).
* Strategies that interleave solely between the 5th and 6th digit are not in this selection.

We can definitely not conclude that humans must favour the "natural breakpoint" between the 5th and 6th digit, since the results for those simulations do not match with human data. However, the data do suggest that humans may take only one or two breaks, where the first break is often after keypress 2, 3, or 4.

### Question 6

We only varied the value of the gaussDeviateSD parameter to compare the use of different drift values on the model fit. 

Because all human-like results came from simulations with only one or two breakpoints, we hypothesised that the number of simulations can be reduced by leaving out the strategies with a very large number of breakpoints. If humans make very few breakpoints, there is no reason to run simulations for a lot of potential strategies that do not match human performance.

To test this hypothesis, we separated the data from question 5 on whether there were at most 5 breaks. These results are plotted below.

```{r q6a_1, echo=FALSE}

less_than_five_breakpoints <- function(strat) {
  return(sum(all_strategies[strat,]) <= 5)
}

complex_sims <- results
complex_sims <- cbind(results, sapply(results$strats, FUN = less_than_five_breakpoints))
colnames(complex_sims) <- c(colnames(complex_sims)[1:5], "at_most_five_breaks")

#plot all strategies

ggplot(data = complex_sims) +
  geom_point(aes(TrialTime, abs(dev), color=at_most_five_breaks), size = 0.2) +
  
  geom_point(aes(x = steer_dial_m, y = steer_dev_m, color = "Steer focus trials")) +
  geom_errorbar(aes(x = steer_dial_m, ymin = steer_dev_m - steer_dev_se, ymax = steer_dev_m + steer_dev_se, color = "Steer focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = steer_dial_m - steer_dial_se, xmax = steer_dial_m + steer_dial_se, y = steer_dev_m, color = "Steer focus trials", height = 0.02)) +
  
  geom_point(aes(x = dial_dial_m, y = dial_dev_m, color =  "Dial focus trials")) +
  geom_errorbar(aes(x = dial_dial_m, ymin = dial_dev_m - dial_dev_se, ymax = dial_dev_m + dial_dev_se, color =  "Dial focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = dial_dial_m - dial_dial_se, xmax = dial_dial_m + dial_dial_se, y = dial_dev_m, color =  "Dial focus trials", height = 0.02)) + 
  
  xlab("Dial time (s)") +
  ylab("Average Lateral Deviation (m)") +
  labs(color = "At most five breakpoints") +
  theme(legend.position = "bottom")

```

As can be seen in the graph, the data from simulations with more than five breakpoints are further away from the human results, and none of them come close to falling within the human range of performance. Based on this, we conclude that it is valid to leave out these strategies. We ran our simulations for different models using all complex strategies where the total number of breakpoints is at most 5.

```{r}

less_than_5_strats <- c()
for (i in 1:nrow(all_strategies)) {
  less_than_5_strats[i] <- sum(all_strategies[i, ]) <= 5
}

strategies_less_than_5 <- all_strategies[less_than_5_strats, ]

# Create 8 models: drift (old/new); IKI (old/new); Nr simulations (10/50)

load("resultsSimpleDriftOldIKIOldSim10.R")
load("resultsSimpleDriftOldIKIOldSim50.R")
load("resultsSimpleDriftOldIKINewSim10.R")
load("resultsSimpleDriftOldIKINewSim50.R")
load("resultsSimpleDriftNewIKIOldSim10.Rdata")
load("resultsSimpleDriftNewIKIOldSim50.Rdata")
load("resultsSimpleDriftNewIKINewSim10.Rdata")
load("resultsSimpleDriftNewIKINewSim50.Rdata")

plotSimulations <- function(results) {

results$TrialTime <- results$TrialTime / 1000

natural_breakpoint_sims <- subset(results, strats == target_strat_n)

#plot all strategies
ggplot(data = results) +
  geom_point(aes(TrialTime, abs(dev), color="Simulations"), size = 0.2) +
  
  geom_point(aes(x = dialTimeSteerMean, y = laneDevSteerMean, color = "Steer focus trials")) +
  geom_errorbar(aes(x = dialTimeSteerMean, ymin = laneDevSteerMean - laneDevSteerSE, ymax = laneDevSteerMean + laneDevSteerSE, color = "Steer focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = dialTimeSteerMean - dialTimeSteerSE, xmax = dialTimeSteerMean + dialTimeSteerSE, y = laneDevSteerMean, color = "Steer focus trials", height = 0.02)) +
  
  geom_point(aes(x = dialTimeDialMean , y = laneDevDialMean, color =  "Dial focus trials")) +
  geom_errorbar(aes(x = dialTimeDialMean , ymin = laneDevDialMean - laneDevDialSE, ymax = laneDevDialMean + laneDevDialSE, color =  "Dial focus trials", width = 0.2)) +
  geom_errorbarh(aes(xmin = dialTimeDialMean  - dialTimeDialSE, xmax = dialTimeDialMean  + dialTimeDialSE, y = laneDevDialMean, color =  "Dial focus trials", height = 0.02)) + 
  
  geom_point(data = natural_breakpoint_sims, aes(TrialTime, abs(dev), color="Natural breakpoint simulations")) +
  
  
  xlab("Dial time (s)") +
  ylab("Average Lateral Deviation (m)") +
  labs(color = "Data source") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.1)) +
  coord_cartesian(ylim = c(0, 1.5))

} 

```

Old drift parameter; old IKI parameter; 10 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftOldIKIOldSim10)
```

Old drift parameter; new IKI parameter; 10 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftOldIKINewSim10)
```

New drift parameter; old IKI parameter; 10 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftNewIKIOldSim10)
```

New drift parameter; new IKI parameter; 10 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftNewIKINewSim10)
```

Old drift parameter; old IKI parameter; 50 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftOldIKIOldSim50)
```


Old drift parameter; new IKI parameter; 50 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftOldIKINewSim50)
```

New drift parameter; old IKI parameter; 50 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftNewIKIOldSim50)
```

New drift parameter; new IKI parameter; 50 simulations

```{r, echo = FALSE, message =FALSE, warning = FALSE}
plotSimulations(resultsSimpleDriftNewIKINewSim50)
```

We will consider the model a good fit of the human data if the results for the steering condition fall within the range of model predictions. For a few of the models, the human data fall completely outside of the range of the model; this means that the model does not predict human-like results for any strategy.
Since we will use the model to draw a conclusion on what strategies humans use (natural breakpoint or not), we should not use predictions for specific strategies to determine the overall model fit. Therefore, we do not look at the natural breakpoint predictions to determine model fit.

Simulations: if we compare the graphs with 10 simulations to those with 50 simulations, we can see that running 10 simulations seems to lead to a slightly better fit with human data. However, this is due to the spread in lateral deviation decreasing with more simulations, because there is more noise in the data. The overall pattern of the model predictions remains the same irregardless of the number of simulations. As argued by Roberts and Pashler (2000), a good model should sufficiently constrain possible outcomes. The 10 simulation predictions have a better fit only because they restrict the possible outcomes less, by allowing more noise. We conclude that the 10 simulation models are not more valid models than the 50 simulation models.

Drift parameter and IKI parameter: for both parameters we see that by calibrating these parameters the model fit increases compared to their old values. In the model where both parameters have their old values, human data seem to be fairly far off from the model predictions. Then, as one of the parameters is calibrated, the model fit increases as the human data fall at the edge of the range of model predictions. Finally, if both parameters are calibrated the human data fall well within the range of the model predictions. 

For these two parameters, we can add that when calibrating them not only do they produce a better fit, but we have a priori motivation to calibrate their values. Since the experiment used a different simulation and a different dialling interface than the original experiment, it makes sense that these parameters should be updated.

Furthermore, the model seems to be quite reliable/generalizable because by adjusting these two parameters, the model still fits the human data well, as in the original experiment. 

For all 8 models, the human data of the steering condition does not overlap with one of the natural breakpoint strategies. As a result, all models indicate that humans do not seem to interleave solely at the natural breakpoint. However, this conclusion can only be drawn under the assumption that the model is valid. This at least requires that some other strategy or strategies should fall within the range of human data, which is not true for all models. 










